Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
WIP

\section{Introduction}
\label{sec:intro}
%
Deep learning \cite{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data? These questions become central when applied to sequential learning paradigms, such as continual learning (CL, \citep{parisi2019continual, de2021continual}) and reinforcement learning (RL, \cite{sutton2018reinforcement}). In CL, access to the previous data is lost, and then the challenge is retaining a compact representation of the problem to alleviate forgetting over the life-long learning horizon~\cite{mccloskey1989catastrophic}. Similarly, in RL, the model must adapt to environmental observations through exploration, while leveraging prediction uncertainties to assess potential future paths.\looseness-1

@@

%Current state of affairs 
Recent techniques (\eg, \cite{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable}) apply the Laplace-GGN approximation\todo{Is Laplace-GGN a well-known acronym?} to convert a trained neural networks into a Bayesian neural network, that can provide uncertainty without sacrificing additional resources to training \cite{foong2019between}. Furthermore, the resultant weight-space posterior can be converted to the function-space as shown in \cite{khan2019approximate, immer2021improving}. The function-space representation allows for a principled mathematical approach for analyzing the behaviour \cite{cho2009kernel,meronen2020stationary}, performing probabilistic inference \cite{khan2019approximate}, and quantifying uncertainty in neural networks \cite{foong2019between}. These methods rely on the linearization of the neural network and the resultant neural tangent kernel (NTK, \cite{jacot2018neural}). The neural network is characterized in function-space by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process (GP, \cite{rasmussen2006gaussian}). GPs provide a widely-used probabilistic toolkit with principled uncertainty estimates. They serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}.  
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

@@

Similarly to the UCI experiments, we trained neural networks (CNN, architecture matches \citet{immer2021improving}) for the FMNIST and CIFAR10 classification tasks. After training, we computed the duals for NN2SVGP and used the resulting SVGP for uncertainty estimates, see \cref{tbl:imagesuper}. We performed hyperparameter optimization over the prior precision, and found the optimal value to be [insert]. We ran the experiment over $5$ seeds. During the NN training, we used a batch size of $512$ and an Adam optimizer with learning rate $1e-3$.

>>>>

\section{Introduction}
\label{sec:intro}
%
Deep learning \cite{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data? These questions become central when applied to sequential learning paradigms, such as continual learning (CL, \citep{parisi2019continual, de2021continual}) and reinforcement learning (RL, \cite{sutton2018reinforcement}). In CL, access to the previous data is lost and the challenge is retaining a compact representation of the problem and avoiding forgetting over the life-long learning horizon~\cite{mccloskey1989catastrophic}. Similarly, in RL, the model must adapt to environmental observations through exploration, while leveraging prediction uncertainties to assess potential future paths.\looseness-1

@@

%Current state of affairs 
Recent techniques (\eg, \cite{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable}) show that converting trained neural networks to a Bayesian neural network through the Laplace-GGN \todo{Is Laplace-GGN a well-known acronym?} can provide uncertainty without sacrificing additional resources to training \cite{foong2019between}. Furthermore, the resultant weight space posterior can be converted to the function space as shown in \cite{khan2019approximate, immer2021improving}. The function space representation allows for a principled mathematical approach for analyzing the behaviour \cite{cho2009kernel,meronen2020stationary}, performing probabilistic inference \cite{khan2019approximate}, and quantifying uncertainty in neural networks \cite{foong2019between}. These methods rely on the linearization of the neural network and the resultant neural tangent kernel (NTK, \cite{jacot2018neural}). The function space neural network is characterized by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process (GP, \cite{rasmussen2006gaussian}). GPs provide a widely-used probabilistic toolkit with principled uncertainty estimates, and they serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}. 
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

@@

Similarly to the UCI experiments, we seek to demonstrate \our on image data. We consider MNIST and Fashion-MNIST (FMNIST), and neural networks (CNN, architecture matches \cite{immer2021improving}) for the FMNIST and CIFAR10 classification tasks. After training, we computed the duals for NN2SVGP and used the resulting SVGP for uncertainty estimates, see \cref{tbl:imagesuper}. We performed hyperparameter optimization over the prior precision, and found the optimal value to be [insert]. We ran the experiment over $5$ seeds. During the NN training, we used a batch size of $512$ and an Adam optimizer with learning rate $1e-3$.
